State of the Practice for Lattice Boltzmann Method Software

### 2022-10-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Submitted version of SOP for LBM
Edits to LBM SOP paper to bring more detail on AHP

\author{Anonymous}
% \author[CAS]{Spencer Smith}
% \author[CAS]{Peter Michalski}
% \author[CAS]{Jacques Carette}
% \author[ME]{Zahra Keshavarz-Motamed}

% \affiliation[CAS]{organization={McMaster University, Computing and Software
% Department}, %Department and Organization
%             addressline={1280 Main Street West}, 
%             city={Hamilton},
%             postcode={L8S 4K1}, 
%             state={Ontario},
%             country={Canada}}

% \affiliation[ME]{organization={McMaster University, Mechanical Engineering}, %Department and Organization
%             addressline={1280 Main Street West}, 
%             city={Hamilton},
%             postcode={L8S 4K1}, 
%             state={Ontario},
%             country={Canada}}

@@

We use 10 research questions to structure and focus our discussion. These
questions drive our methodology. 
% \citep{SmithEtAl2021, SmithAndMichalski2022} % DOUBLE BLIND
For each question below, we point to the section that contains our answer.  We
start with identifying the examples of LBM software where we have access to the
source code:

@@

To understand the state of the practice, we wish to learn the frequency with
which different artifacts appear, the types of development tools used, and the
methodologies used for software development.  With this data, we can ask
questions about how LBM software compares to other research software.  For cases
where data is available, we can compare to other specific domains of research
software, like medical imaging \citep{Dong2021} and ocean modelling
\citep{JungEtAl2022}.  However, most of the literature on the state of the
practice does not focus on specific domains, but rather on all research software.
%\citep{SmithAndMichalski2022}. % DOUBLE BLIND
We will refer to studies and guidelines that do
not distinguish an application domain as \emph{research software in general}. By
comparing LBM software to research software in general we can determine how LBM
follows the trends from other developer communities and how it is different.

@@

We investigate the research questions by applying the general methodology
summarized in Anonymous (xxxx)\footnote{Anonymous for double-blind
review}.  %\citet{SmithEtAl2021, SmithAndMichalski2022}.  %DOUBLE BLIND
The specific
application of the methodology to LBM is reviewed in Section~\ref{methodology},
with the full details in Anonymous (xxxx)\footnotemark[1].  %\citet{Michalski2021}.  
Our methodology updates
an approach used in prior assessments of domains like Geographic Information
Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Oceanographic Software \citep{smith2015state}, Seismology software
\citep{SmithEtAl2018}, statistical software for psychology
\citep{SmithEtAl2018_StatSoft} and medical image analysis software
\citep{Dong2021}.

@@

Compared with our previous methodology (Anonymous, xxxx)\footnotemark[1],
%(as used in \citet{SmithEtAl2016} for instance) %DOUBLE BLIND
measuring the application of best practices now also includes
repository based metrics, such as the number of files, number of lines of code,
percentage of issues that are closed, etc.  We rank the software with the
Analytic Hierarchy Process (AHP), using the quantitative data from the grading
template. After this, as another addition to our previous methodology, we
interview some development teams to understand the status of their
development processes. We next answer the research questions using the collected
quantitative and qualitative data (Sections~\ref{methodology}
to~\ref{Sec_Recommendations}).  Finally, we summarize the threats to validity
(Section~\ref{threats}) and the final conclusions and recommendations
(Section~\ref{conclusion}).

@@

We apply our methodology for evaluating the state of the practice of research
software %\citep{SmithEtAl2021, SmithAndMichalski2022} % DOUBLE BLIND
to the specific domain of
LBM software.  Since we group best practices around conventional software
qualities, the first section below provides definitions for the qualities of
interest (Section~\ref{softwarequalities}), along with information on how to
measure these qualities. The following section
(Section~\ref{Sec_OverallProcess}) provides an overview of the steps we used to
select, measure and compare LBM software.  The remaining sections
(Sections~\ref{identifysoftware} to~\ref{sec_vet_software_list}) provide details
on each step in the methodology.

@@

\begin{enumerate}
	\item List candidate software packages for the domain.
	(Section~\ref{identifysoftware})
	\item Filter the software package list. (Section~\ref{filtersoftware})
	\item Gather the source code and documentation for each software package.
	\item Collect quantitative measures from the project repositories.
	(Section~\ref{empiricalmeasures})
	\item Measure using the measurement template.  The full measurement template
	can be found in Anonymous (xxxx)\footnotemark[1]. %\citet{SmithEtAl2021}. %DOUBLE BLIND
	(Section~\ref{empiricalmeasures})
	\item Rank the software using AHP. (Section~\ref{AHP})
	\item Interview the developers. (Section~\ref{SecSurvey})
	\item Analyze the domain. (Section~\ref{SecDomainAnalysis})
	\item Analyze the results and answer the research questions.
\end{enumerate}

@@

We use a measurement template to rank the projects by how well they follow best
practices (\rqref{RQ_HighestQuality}), as described in Anonymous (xxxx)\footnotemark[1].
% \citet{SmithEtAl2021}. % DOUBLE BLIND
For each software package (each column in the template), we fill in the rows of
the template. This process takes about two hours per package, with a cap of four
hours. The time constraint is necessary so that the workload is feasible for a
team as small as one, given our aim to cap the measurement phase at 160
person-hours (Anonymous, xxxx)\footnotemark[1]. %\citep{SmithEtAl2021}. %DOUBLE BLIND
An excerpt of the template, in spreadsheet
form, is shown in Figure~\ref{measurement_template_image}.

@@

%As in \citet{SmithEtAl2016}, %DOUBLE BLIND
We used Virtual Machines (VMs) to provide an
optimal testing environment for each package. We used VMs to create fresh
environments, eliminating any worries about existing libraries and conflicts.
Moreover, when the tests are complete, the VM can be deleted, without impacting
the host operating system. The most significant advantage of using VMs is that
they level the playing field. Every software install starts from a clean slate,
which removes ``works-on-my-computer'' errors.

@@

Once we measure each package, we still need to rank them to answer
\rqref{RQ_HighestQuality}. We do this via the Analytic Hierarchy Process (AHP),
a decision-making technique that compares multiple options by multiple criteria.
Our AHP algorithm uses the overall impression score, gathered via the
measurement template, to compare and rank the LBM software. AHP performs a
pairwise analysis using a matrix and generates an individual quality score for
each software package. Anonymous (xxxx)\footnotemark[1] %\citet{SmithEtAl2016} %DOUBLE BLIND
shows the details on how AHP ranks
software based on quality measures.

We use our own simple tool for conducting the AHP analysis. The tool includes a
sensitivity analysis to ensure that the software package rankings are
appropriate with respect to the uncertainty of the quality scores. For the
sensitivity analysis, we modified the score by 10\% for each package and
verified that the overall ranking was stable.  
% The
% \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{README}
% file for the tool includes requirements and usage information. %DOUBLE BLIND

@@

Two of the research question (\rqref{RQ_PainPoints} and \rqref{RQ_Concerns})
require going beyond the quantitative data from the measurement template. To
gain insight, we interview developers using a list of 20 questions from Anonymous (xxxx)\footnotemark[1].
%\citet{SmithEtAl2021}. %DOUBLE BLIND
The questions cover the background of the development
teams, the interviewees, and the software itself. We ask the developers how they
organize their projects and about their understanding of the users. Some
questions focus on the current and past difficulties, and the solutions the team
has found or will try. We also discuss the importance of, and the current
situation for, documentation. A few questions focus on specific software
qualities, such as maintainability, understandability, usability, and
reproducibility. The interviews are semi-structured based on the question list;
we ask follow-up questions when necessary. Based on our experience, the
interviewees often raise exciting and unexpected ideas.

We requested interviews with developers for all 24 packages, except Musubi,
since we added Musubi after our ethics approval expired. We found the developers
from project websites, code repositories, publications, and biographic
information on institution webpages. Four developers agreed to participate in
our study. We met with them online, using either Zoom or Teams to facilitate
recording and automatic transcription. Each meeting lasted about an hour. 
% The
% McMaster University Research Ethics Board approved the interview process under
% application number
% \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
% {MREB\#: 5219}. %DOUBLE BLIND

@@

We answer \rqref{RQ_HighestQuality} by using AHP to rank the identified LBM
software against best practices.  The sections below summarize how the software
packages compare based on the qualities discussed in
Section~\ref{softwarequalities}, as measured via the measurement template
(Section~\ref{empiricalmeasures}).  The last subsection combines all the
qualities into one ``global'' ranking.  The full LBM data is available publicly
on Mendeley Data (Anonymous, xxxx)\footnotemark[1]. %\citep{Smith2022}. %DOUBLE BLIND

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking. In the absence of other
information on priorities, we calculated the ranking by assuming an equal
weighting between all qualities. Software packages with an overall high score
ranked high in at least several of the individual qualities.

@@

Table~\ref{Tbl_Guidelines} %(based on data from \citep{SmithAndMichalski2022}) %DOUBLE BLIND
shows that MI artifacts generally match the recommendations found in nine
current research software development guidelines:
\begin{itemize}
\item United States Geological Survey Software Planning Checklist
\citep{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\citep{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \citep{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \citep{SmithAndRoscoe2018},
\item Trilinos Developers Guide \citep{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \citep{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \citep{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \citep{OrvizEtAl2017}.
\end{itemize}

@@

We investigated the risk of reviewer dependence in previous work (Anonymous, xxxx)\footnotemark[1],
%\citep{SmithEtAl2016} % DOUBLE BLIND
where we demonstrated that the measurement process is
reasonably reproducible by having five products graded by a second reviewer. The
ranking via this independent review was almost identical to the original
ranking. As long as each grader uses consistent definitions, the relative
comparisons in the AHP results will be consistent between graders.

@@

>>>>

%\author{Anonymous}
\author[CAS]{Spencer Smith}
\author[CAS]{Peter Michalski}
\author[CAS]{Jacques Carette}
\author[ME]{Zahra Keshavarz-Motamed}

\affiliation[CAS]{organization={McMaster University, Computing and Software
Department}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\affiliation[ME]{organization={McMaster University, Mechanical Engineering}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

@@

We use 10 research questions to structure and focus our discussion. These
questions drive our methodology \citep{SmithEtAl2021, SmithAndMichalski2022}.
For each question below, we point to the section that contains our answer.  We
start with identifying the examples of LBM software where we have access to the
source code:

@@

To understand the state of the practice, we wish to learn the frequency with
which different artifacts appear, the types of development tools used, and the
methodologies used for software development.  With this data, we can ask
questions about how LBM software compares to other research software.  For cases
where data is available, we can compare to other specific domains of research
software, like medical imaging \citep{Dong2021} and ocean modelling
\citep{JungEtAl2022}.  However, most of the literature on the state of the
practice does not focus on specific domains, but rather on all research software
\citep{SmithAndMichalski2022}. We will refer to studies and guidelines that do
not distinguish an application domain as \emph{research software in general}. By
comparing LBM software to research software in general we can determine how LBM
follows the trends from other developer communities and how it is different.

@@

We investigate the research questions by applying the general methodology
summarized in \citet{SmithEtAl2021, SmithAndMichalski2022}. The specific
application of the methodology to LBM is reviewed in Section~\ref{methodology},
with the full details in Anonymous (xxxx)\footnotemark[1].
%\citet{Michalski2021}.  
Our methodology updates
an approach used in prior assessments of domains like Geographic Information
Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Oceanographic Software \citep{smith2015state}, Seismology software
\citep{SmithEtAl2018}, statistical software for psychology
\citep{SmithEtAl2018_StatSoft} and medical image analysis software
\citep{Dong2021}.

@@

Compared with our previous methodology (as used in \citet{SmithEtAl2016} for
instance) %(Anonymous, xxxx)\footnotemark[1] %DOUBLE BLIND
measuring the application of best practices now also includes repository based
metrics, such as the number of files, number of lines of code, percentage of
issues that are closed, etc.  We rank the software with the Analytic Hierarchy
Process (AHP), using the quantitative data from the grading template. After
this, as another addition to our previous methodology, we interview some
development teams to understand the status of their development processes. We
next answer the research questions using the collected quantitative and
qualitative data (Sections~\ref{methodology} to~\ref{Sec_Recommendations}).
Finally, we summarize the threats to validity (Section~\ref{threats}) and the
final conclusions and recommendations (Section~\ref{conclusion}).

@@

We apply our methodology for evaluating the state of the practice of research
software \citep{SmithEtAl2021, SmithAndMichalski2022} % DOUBLE BLIND
to the specific domain of
LBM software.  Since we group best practices around conventional software
qualities, the first section below provides definitions for the qualities of
interest (Section~\ref{softwarequalities}), along with information on how to
measure these qualities. The following section
(Section~\ref{Sec_OverallProcess}) provides an overview of the steps we used to
select, measure and compare LBM software.  The remaining sections
(Sections~\ref{identifysoftware} to~\ref{sec_vet_software_list}) provide details
on each step in the methodology.

@@

\begin{enumerate}
	\item List candidate software packages for the domain.
	(Section~\ref{identifysoftware})
	\item Filter the software package list. (Section~\ref{filtersoftware})
	\item Gather the source code and documentation for each software package.
	\item Collect quantitative measures from the project repositories.
	(Section~\ref{empiricalmeasures})
	\item Measure using the measurement template.  The full measurement template
	can be found in \citet{SmithEtAl2021}. %DOUBLE BLIND
	(Section~\ref{empiricalmeasures})
	\item Rank the software using AHP. (Section~\ref{AHP})
	\item Interview the developers. (Section~\ref{SecSurvey})
	\item Analyze the domain. (Section~\ref{SecDomainAnalysis})
	\item Analyze the results and answer the research questions.
\end{enumerate}

@@

We use a measurement template to rank the projects by how well they follow best
practices (\rqref{RQ_HighestQuality}), as described in \citet{SmithEtAl2021}. % DOUBLE BLIND
For each software package (each column in the template), we fill in the rows of
the template. This process takes about two hours per package, with a cap of four
hours. The time constraint is necessary so that the workload is feasible for a
team as small as one, given our aim to cap the measurement phase at 160
person-hours \citep{SmithEtAl2021}. %DOUBLE BLIND
An excerpt of the template, in spreadsheet
form, is shown in Figure~\ref{measurement_template_image}.

@@

As in \citet{SmithEtAl2016}, %DOUBLE BLIND
we used Virtual Machines (VMs) to provide an
optimal testing environment for each package. We used VMs to create fresh
environments, eliminating any worries about existing libraries and conflicts.
Moreover, when the tests are complete, the VM can be deleted, without impacting
the host operating system. The most significant advantage of using VMs is that
they level the playing field. Every software install starts from a clean slate,
which removes ``works-on-my-computer'' errors.

@@

Once we measure each package, we still need to rank them to answer
\rqref{RQ_HighestQuality}. We do this via the Analytic Hierarchy Process (AHP),
a decision-making technique that compares multiple options by multiple criteria.
Our AHP algorithm uses the overall impression score, gathered via the
measurement template, to compare and rank the LBM software. AHP performs a
pairwise analysis using a matrix and generates an individual quality score for
each software package.  We have previously used this technique to rank software
in different domains, as in \citet{SmithEtAl2016}.

AHP focuses on relative comparisons, rather than requiring an unattainable
unified scale for measuring quality.  AHP starts with sets of $n$
\textit{options} and $m$ \textit{criteria}.  In our project there are 24
software products ($n=24$) and 9 criteria ($m = 9$).  To get a ranking for each
criterion, we perform a pairwise analysis between each of the options, in the
form of an $n \times n$ matrix $A$.  The value of $A_{jk}$ ranges from 1 to 9,
using the following definitions \citep{Saaty1980}:

\begin{itemize}
\item 1 means equal importance. Two activities contribute equally to the
  objective.
\item 3 means weak importance of one over another. Experience and judgment
  slightly favor one activity over another.
\item 5 means essential or strong importance. Experience and judgment strongly
  favor one activity over another.
\item 7 means demonstrated importance. An activity is strongly favored and its
  dominance demonstrated in practice.
\item 9 means absolute importance. The evidence favoring one activity over
  another is of the highest possible order of affirmation.
\item 2, 4, 6, 8 are defined as intermediate values between the two adjacent
  judgments.
\end{itemize}

During our grading process we capture a subjective score, from one to ten, for
each package for each criterion.  To turn these into Saaty's pair-wise scores
for option $x$ versus option $y$, we use the following calculation for the
respective subjective scores $x_{\text{sub}}$ and $y_{\text{sub}}$:
\[
\begin{cases}
\min\{9, x_{\text{sub}} - y_{\text{sub}} + 1\} & x_{\text{sub}} \geq y_{\text{sub}} \\
1 / \min\{9, y_{\text{sub}} - x_{\text{sub}} + 1\} & x_{\text{sub}} < y_{\text{sub}}
\end{cases}
\]

For example, the installability for DL\_MESO and waLBerla is measured as $9$ and
$4$, respectively, so the score for DL\_MESO compared to waLBerla is 6 and
waLBerla compared to DL\_MESO is 1/6, as shown in Figure~\ref{}.

In our work, we make the common assumption of a reciprocal relationship between
options $k$ and $j$; that is, $A_{kj}=1/A_{jk}$.  A partial view of $A$ for the
quality of installability is found in Figure~\ref{}.  To obtain the overall
score for each option for a given criterion we use matrix $B$, where $B_{jk} =
A_{jk}/\sum_{i=1}^{i=n}(A_{i k})$.  That is, to get $B$ we normalize the columns
of $A$, as shown in Figure~\ref{}.  For a given quality, the score ($w_j$) for
software package $j$ is found by averaging the entries in row $j$ of $B$.  That
is, $w_j = \frac{1}{m} \sum_{i=1}^{i=m}(A_{j i})$. Besides providing a ranking,
the $w_j$ values can be combined for all qualities for an overall ranking, as
discussed in Section~\ref{Sec_OverallQuality}.

Figure

We use our own simple tool for conducting the AHP analysis. The tool includes a
sensitivity analysis to ensure that the software package rankings are
appropriate with respect to the uncertainty of the quality scores. For the
sensitivity analysis, we modified the score by 10\% for each package and
verified that the overall ranking was stable.  
The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{README}
file for the tool includes requirements and usage information. %DOUBLE BLIND

@@

Two of the research question (\rqref{RQ_PainPoints} and \rqref{RQ_Concerns})
require going beyond the quantitative data from the measurement template. To
gain insight, we interview developers using a list of 20 questions from 
\citet{SmithEtAl2021}. %DOUBLE BLIND
The questions cover the background of the development
teams, the interviewees, and the software itself. We ask the developers how they
organize their projects and about their understanding of the users. Some
questions focus on the current and past difficulties, and the solutions the team
has found or will try. We also discuss the importance of, and the current
situation for, documentation. A few questions focus on specific software
qualities, such as maintainability, understandability, usability, and
reproducibility. The interviews are semi-structured based on the question list;
we ask follow-up questions when necessary. Based on our experience, the
interviewees often raise exciting and unexpected ideas.

We requested interviews with developers for all 24 packages, except Musubi,
since we added Musubi after our ethics approval expired. We found the developers
from project websites, code repositories, publications, and biographic
information on institution webpages. Four developers agreed to participate in
our study. We met with them online, using either Zoom or Teams to facilitate
recording and automatic transcription. Each meeting lasted about an hour. 
The
McMaster University Research Ethics Board approved the interview process under
application number
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}. %DOUBLE BLIND

@@

We answer \rqref{RQ_HighestQuality} by using AHP to rank the identified LBM
software against best practices.  The sections below summarize how the software
packages compare based on the qualities discussed in
Section~\ref{softwarequalities}, as measured via the measurement template
(Section~\ref{empiricalmeasures}).  The last subsection combines all the
qualities into one ``global'' ranking.  The full LBM data is available publicly
on Mendeley Data \citep{Smith2022}. %DOUBLE BLIND

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking. In the absence of other
information on priorities, we calculated the ranking by assuming an equal
weighting between all qualities. Software packages with an overall high score
ranked high in at least several of the individual qualities. ****TODO: Explain how
to combine rankings.****

@@

Table~\ref{Tbl_Guidelines} (based on data from \citep{SmithAndMichalski2022}) %DOUBLE BLIND
shows that LBM artifacts generally match the recommendations found in nine
current research software development guidelines:
\begin{itemize}
\item United States Geological Survey Software Planning Checklist
\citep{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\citep{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \citep{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \citep{SmithAndRoscoe2018},
\item Trilinos Developers Guide \citep{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \citep{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \citep{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \citep{OrvizEtAl2017}.
\end{itemize}

@@

We investigated the risk of reviewer dependence in previous work \citep{SmithEtAl2016}, % DOUBLE BLIND
where we demonstrated that the measurement process is
reasonably reproducible by having five products graded by a second reviewer. The
ranking via this independent review was almost identical to the original
ranking. As long as each grader uses consistent definitions, the relative
comparisons in the AHP results will be consistent between graders.

@@

\section*{Conflict of Interest}

On behalf of all authors, the corresponding author states that there is no
conflict of interest.
